A Dearth of Young Stars and Planets in the Kepler Field

### 2024-05-27 paper/ms.tex
update ms

Using our effective temperatures and rotation periods, we ran
\texttt{gyro-interp} over an age grid of 0 to 4\,Gyr, linearly spaced
over 500 grid points.  \texttt{gyro-interp} itself is calibrated out
to 4\,Gyr, the age of M67
\citep[see][]{Dungee_2022,Gruner_2023}.  To estimate the
stellar spin-down rate as a function of time and stellar temperature,
the framework interpolates between reference open clusters using cubic
Hermite polynomials.  We calculate the probability of the
rotation-based age $t_{\rm gyro}$, given the observed rotation
periods, effective temperatures, and their uncertainties by
integrating Equation~1 of \citet{Bouma_2023} at the default grid
resolution.  This procedure marginalizes over the intrinsic age and
temperature dependent scatter that is observed in the cluster
sequences to yield statistical age uncertainties.

@@

{\it Grazing planets (bit 2)}---Grazing objects, for which the impact
parameter $b$ is greater than $1-R_{\rm p}/R_\star$, often yield
biased planetary parameters \citep[e.g.][]{2022AJ....163..111G}.  For
large planets, they also include astrophysical false positives at
higher rates \citep{2016ApJ...822...86M}, in part due to the
size-impact parameter degeneracy.  We flag planet candidates as
potentially grazing if $b>0.8$, using the impact parameters reported
in the Markov Chain Monte Carlo (MCMC) fits from
\citet{Thompson_2018}.

@@

In Figure~\ref{fig:agescalecompone}, we compare our rotation-based
ages, as well as select ages from the literature, against the ages of
stars in the richest clusters in the Kepler field.  From the
literature, we drew ages from \citet{Reinhold_2015},
\citet{Berger_2020a_catalog}, \citet{2021AJ....161..189L},
\citet{2023ApJ...952..131M}, and \citet{2024AJ....167..159L}.  We
followed any guidance available from these respective studies for
removing ages that were explicitly flagged as unreliable, and plotted
stars within $1\sigma$ of zero as being upper limits.  For the
\citeauthor{Reinhold_2015} ages, we used those that were calculated
using the \citet{Mamajek_2008} calibration; for the
\citeauthor{2023ApJ...952..131M} ages, we show those from
\texttt{kiauhoku} \citep{Claytor2020} rather than \texttt{STAREVOL}
\citep{Amard2019}, based on the defaults adopted by
\citeauthor{2023ApJ...952..131M}.

Cluster membership is a delicate subject.
Figure~\ref{fig:agescalecompone} is showing reported ages for a set of
spatially and kinematically selected set of stars that could be
cluster members, or they could be field interlopers.  For NGC6811 and
NGC6819, we adopted candidate members from
\citet{2018A&A...618A..93C,CantatGaudin_2020} and
\citet{Kounkel_2020}.  For Melange-3 we used the candidates reported
by \citet{Barber_2022} and required ``offset'' tangential velocities
below 2\,\kms.  For Cep-Her we used candidate members from Table~2 of
\citet{Bouma_2022b}, requiring ``weights'' exceeding 0.1.  Even with
contaminants, we can compare the relative age distributions derived by
the different studies because in all instances we are comparing
reported ages against a fixed list of stars.

There are two main metrics for success in this test: accuracy in
whether the reported ages agree with the cluster age, and precision in
whether the reported age uncertainties agree with their dispersion
around the cluster age.

The point of Figure~\ref{fig:agescalecompone} is hopefully clear.
While previous studies have been able to derive accurate ages for
stars $\gtrsim$1\,Gyr old, the ages of sub-gigayear stars have
historically been overestimated by anywhere from 0.3--2\,dex.  The
situation is particularly severe for the 40\,Myr stars in Cep-Her, but
it is also present for 150--300\,Myr stars in Melange-3 and Theia~520.
For isochrone and kinematic ages, this is not surprising because these
methods rely on parameters that do not appreciably change at
$t$$\lesssim$1\,Gyr.  For the rotation-based ages quoted by
\citet{2023ApJ...952..131M} and \citet{2024AJ....167..159L}, the
discrepancy is likely
caused by details of the calibration methodologies.  The physical
spin-down models used by \citet{2023ApJ...952..131M}, and the
gyro-kinematic calibration ages used by \citet{2024AJ....167..159L},
may not reproduce the rotation sequences of open clusters such as
Praesepe, the Hyades, and the Pleiades.  The \citet{Mamajek_2008}
calibration used by \citet{Reinhold_2015} does reproduce such
sequences for Sun-like stars.  However the
uncertainties determined using that method are underestimates because
they do not marginalize over
the intrinsic scatter in rotation periods present for stars of a fixed
mass and age.

@@

Appendix~\ref{app:inconsistent} describes all cases of ``discrepant''
confirmed planets for which a sub-gigayear age is reported by at least
one age indicator.  All but two systems are either evolved stars or
else unresolved binaries, and are automatically flagged as such.  The
more interesting of the two is Kepler-786,
an early K dwarf with $t_{\rm
Li}$=\kepseveneightsix\,Myr, which would imply that rotation should be
detectable.  However no photometric rotation signal is detected; other
age indicators in the spectrum similarly suggest that the star may not
in fact be young.

>>>>

To estimate the
stellar spin-down rate as a function of time and stellar temperature,
\texttt{gyro-interp} uses measured rotation periods and effective
temperatures from reference open clusters, and interpolates between them using cubic
Hermite polynomials.  We calculated the probability of the
rotation-based age $t_{\rm gyro}$, given the observed rotation
periods and effective temperatures (and their uncertainties) by
integrating Equation~1 of \citet{Bouma_2023} at the default
resolution.  This procedure marginalizes over the intrinsic age and
temperature dependent scatter that is observed in the cluster
sequences to yield statistical age uncertainties.
We adopted a linear age grid spanning 0--5\,Gyr with
500 grid points, and interpolated the resulting posteriors to determine
point estimate summary statistics.
At the old end, 
\texttt{gyro-interp} is calibrated out
to 4\,Gyr M67 cluster
\citep[see][]{Dungee_2022,Gruner_2023}.  
For Kepler stars with rotation periods above the 
M67 sequence, our resulting age constraints are quoted as lower limits.

@@

{\it Grazing planets (bit 2)}---Grazing objects, for which the impact
parameter $b$ is greater than $1-R_{\rm p}/R_\star$, often yield
biased planetary parameters \citep[e.g.][]{2022AJ....163..111G}.  For
large planets, they also include astrophysical false positives at
higher rates \citep{2016ApJ...822...86M}, in part due to the
size-impact parameter degeneracy.  We flag planet candidates as
potentially grazing if $b$$>$0.8, using the impact parameters reported
in the Markov Chain Monte Carlo (MCMC) fits from
\citet{Thompson_2018}.

@@

In Figure~\ref{fig:agescalecompone}, we compare our rotation-based
ages, and available ages from the literature, against the ages of
stars in open clusters in the Kepler field.  From the
literature, we drew ages from \citet{Reinhold_2015},
\citet{Berger_2020a_catalog}, \citet{2021AJ....161..189L},
\citet{2023ApJ...952..131M}, and \citet{2024AJ....167..159L}.  We
followed any guidance available from these respective studies for
removing ages that were explicitly flagged as unreliable, and plotted
stars within $1\sigma$ of zero upper limits.  For the
\citeauthor{Reinhold_2015} ages, we used those calculated
using the \citet{Mamajek_2008} calibration; for the
\citeauthor{2023ApJ...952..131M} ages, we show those from
\texttt{kiauhoku} \citep{Claytor2020} rather than \texttt{STAREVOL}
\citep{Amard2019}, since it showed better agreement with
the cluster age scale.

Cluster membership is a nuanced subject.
Figure~\ref{fig:agescalecompone} is showing reported ages for a set of
spatially and kinematically selected stars that could be
cluster members, or they could be field interlopers.  For NGC6811 and
NGC6819, we adopted candidate members from
\citet{2018A&A...618A..93C,CantatGaudin_2020} and
\citet{Kounkel_2020}.  For Melange-3 we used the candidates reported
by \citet{Barber_2022} and required ``offset'' tangential velocities
below 2\,\kms.  For Cep-Her we used candidate members from Table~2 of
\citet{Bouma_2022b} with ``weights'' exceeding 0.1.  Even with
contaminants, we can compare the relative age distributions derived by
the different studies because in all instances we are comparing
reported ages against a fixed list of stars.

There are two main metrics for success in this test: {\it 1)}
do the reported ages agree with the cluster age? And {\it 2)} do the
reported age uncertainties agree with their dispersion
around the cluster age?

The point of Figure~\ref{fig:agescalecompone} is hopefully clear.
While previous studies have been able to derive accurate ages for
stars $\gtrsim$1\,Gyr old, available literature ages of sub-gigayear stars are
usually overestimated by 0.3--2\,dex.  
The ages of the youngest Kepler stars and planets were not previously well-measured.
For isochrone and kinematic ages, this is because these
methods rely on parameters that do not appreciably change at
$t$$\lesssim$1\,Gyr.  For the rotation-based ages from
\citet{2023ApJ...952..131M} and \citet{2024AJ....167..159L}, the
discrepancy is
caused by details of the calibration methodologies.  The
spin-down models used in these studies
may struggle to reproduce the observed rotation
sequences seen in open clusters younger than a few hundred million years.  The \citet{Mamajek_2008}
calibration used by \citet{Reinhold_2015} does fit such
sequences for Sun-like stars.  However the
uncertainties from all of these methods are underestimated because
they do not marginalize over
the range of rotation periods that stars of a fixed
mass and age might have.

@@

Appendix~\ref{app:inconsistent} describes all cases of ``discrepant''
confirmed planets for which a sub-gigayear age is reported by at least
one age indicator.  All but two systems are either evolved stars or
else unresolved binaries, and are automatically flagged as such.  The
more interesting of the two is Kepler-786,
an early K dwarf with $t_{\rm
Li}$=\kepseveneightsix\,Myr, which would imply that rotation should be
detectable.  However no photometric rotation signal is detected, and other
age indicators in the spectrum similarly suggest that the star is not
young.
